EuPPollNet: A European database of plant-pollinator networks

### 2024-10-07 Manuscript/EuPPollNet_main_text.qmd
minor submission updates

**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants. Indeed, the network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning. However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interaction networks. One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time. Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest. This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences. We present the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks. We hope EuPPollNet will stimulate future research that fills the taxonomic, ecological, and geographical data gaps on plant-pollinator interactions that we have identified. Further, the variation in the structure of the networks in EuPPollNet provides a strong basis for future studies aimed at quantifying drivers of plant-pollinator network change and guiding future conservation planning for plants and pollinators.

**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries. In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., genus, family and order).

@@

**Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`. All records are time-referenced and most of the studies documented interactions within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).

@@

**Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats. The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.

@@

All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).

>>>>

**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity. However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interactions.  Unfortunately, we lack robust generalisable data capturing how plant-pollinator communities are structured across space and time. Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest. In addition, we evaluate the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks. We believe EuPPollNet will stimulate research to address data gaps in plant-pollinator interactions and guide future efforts in conservation planning.

**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks, which belong to `r study_number` different studies distributed across `r country_number` European countries. Information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information are also provided.

@@

**Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`.

@@

**Software format:** The database was built with R and is stored in “.rds” and “.csv” formats. Its construction is fully reproducible and can be accessed at: <https://doi.org/10.5281/zenodo.13897566>.

@@

All data and code to produce of this database and manuscript are available at Zenodo (<https://doi.org/10.5281/zenodo.13897566>).

### 2024-10-07 Manuscript/Title_page.qmd
Zenodo link
minor submission updates

Jose B. Lanuza is a former master's, PhD, and postdoctoral researcher under Dr. Ignasi Bartomeus. Their work focuses on understanding the ecological processes that drive plant-pollinator interactions at different ecological scales. This research is another piece of the puzzle that will contribute to our understanding of how interactions change with various environmental drivers. Our work has primarily involved flowering plants and pollinators, where we have addressed questions related to their coexistence, interaction probability, and ecology. We have created the first database of Iberian bees for Spain and are strong advocates for reproducible science; this work is a perfect example of it. Further examples of the ongoing work and past projects can be consulted at Ignasi Bartomeus lab web page: <https://bartomeuslab.com/projects/>.

>>>>

Dr. Jose B. Lanuza is a long term collaborator of Dr. Ignasi Bartomeus. Their work focuses on understanding the ecological processes that drive plant-pollinator interactions at different ecological scales. This research is another piece of the puzzle that will contribute to our understanding of how interactions change with various environmental drivers. Their work has primarily involved flowering plants and pollinators, where they have addressed questions related to their coexistence, interaction probability, and ecology. They have created the first database of Iberian bees for Spain and are strong advocates for reproducible science; this work is a perfect example of it. Further examples of the ongoing work and past projects can be consulted at Ignasi Bartomeus lab web page: <https://bartomeuslab.com/projects/>.
